The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-09-26 article.tex
Reverted to the main style, as the others are too hard to read
Reworked abstract Fixed width spec of multirow element

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
%\documentclass[10pt]{bmcart}
\documentclass[doublespacing,linenumbers]{bmcart}
%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

@@

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
Cheminformatics is a well-established field with many applications in chemistry,
biology, drug discovery, and others. The Chemistry Development Kit (CDK) has
become a widely used Open Source cheminformatics toolkit, providing various models to represent
chemical structures, of which the chemical graph is essential. However,
in the first five years of the project increased so much in size that
interdependencies between components grew unmanageable large, resulting
in unpredictable instabilities.
\parttitle{Results}
We here report improvements to the CDK since the 1.2 release series made
to accommodate both the increased complexity of the library, as well as
significant improvements of and additions to the functionality of the
library. Second, we outline how the CDK evolved with respect to
quality control and the approach we have adopted to ensure stability, including
a peer review mechanism.
Additionally, a selection of the new APIs that have been introduced will
be discussed: atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
\parttitle{Conclusions}
With this paper we have shown the continued effort to provide a free, Open Source
cheminformatics library, and show that such collaborative projects can exist over
a long period and result in a high-quality and fast library. We have taken
advantage from the community support, and show
that an open source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

@@

    \begin{minipage}{1\textwidth}
    \centering
    \begin{tabular}{llllllll}
	\multirow{2}{*}{ Input } & \multirow{2}{*}{ Mass tolerance ($\pm$ Da) } & \multicolumn{3}{c}{ \# of generated formulas } & \multicolumn{3}{c}{ Runtime (s) } \\
	& & HR2 & PFG & CDK & HR2 & PFG & CDK \\
	$10\,000$ small masses & $0.001$ & $616\,846$ & $616\,846$ & $616\,843$ & $669$ & $168$ & $\mathbf{41}$ \\
	$10\,000$ small masses & $0.01$ & $6\,163\,303$ & $6\,163\,302$ & $6\,163\,326$ & $689$ & $501$ & $\mathbf{212}$ \\
	$20$ large masses & $0.001$ & $4\,912\,939$ & $4\,912\,939$ & $4\,912\,904$ & $26\,370$ & $1\,292$ & $\mathbf{177}$ \\
	$20$ large masses & $0.01$ & $49\,128\,811$ & $49\,128\,810$ & $49\,128\,815$ & $26\,587$ & $3\,406$ & $\mathbf{1\,580}$ \\
    \end{tabular}
    \end{minipage}

>>>>

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmcart}
%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

@@

\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.
\parttitle{Background}
The Chemistry Development Kit (CDK) is a 
widely used Open Source cheminformatics toolkit, providing data
structures to represent chemical concepts along with methods to
manipulate such structures and perform computations on them. The
library implements a wide variety of cheminformatics algorithms
ranging from chemical structure canonicalization to molecular
descriptor calculations and pharmacophore perception. Over the
last five years, the code based has grown significantly resulting in
many, complex interdependencies between components and poor
performance of many algorithms. 
\parttitle{Results}
We report improvements to the CDK since the 1.2 release series,
specifically addressing the increase functional complexity and poor
performance. We first summarize the addition of new, and
improvements to existing functionality that has led to significant
improvement in performance and support for important chemistry
features including atom type perception, substructure searching, molecular
fingerprints, rendering of molecules, and handling of molecular formulas.
Second, we outline how the CDK has evolved with respect to
quality control and the approaches we have adopted to ensure stability, including
a code review mechanism.
\parttitle{Conclusions}
This paper highlights the our continued efforts to provide a community
driven, Open Source
cheminformatics library, and show that such collaborative projects can exist over
extended periods of time, resulting in a high-quality and performant
library. By taking 
advantage of community support and contributions, we show 
that an Open Source cheminformatics project can act as a peer reviewed
publishing platform for scientific computing software.
\end{abstract}

@@

    \begin{minipage}{1\textwidth}
    \centering
    \begin{tabular}{llllllll}
	\multirow{2}{1in}{ Input } & \multirow{2}{1in}{ Mass tolerance ($\pm$ Da) } & \multicolumn{3}{c}{ \# of generated formulas } & \multicolumn{3}{c}{ Runtime (s) } \\
	& & HR2 & PFG & CDK & HR2 & PFG & CDK \\
	$10\,000$ small masses & $0.001$ & $616\,846$ & $616\,846$ & $616\,843$ & $669$ & $168$ & $\mathbf{41}$ \\
	$10\,000$ small masses & $0.01$ & $6\,163\,303$ & $6\,163\,302$ & $6\,163\,326$ & $689$ & $501$ & $\mathbf{212}$ \\
	$20$ large masses & $0.001$ & $4\,912\,939$ & $4\,912\,939$ & $4\,912\,904$ & $26\,370$ & $1\,292$ & $\mathbf{177}$ \\
	$20$ large masses & $0.01$ & $49\,128\,811$ & $49\,128\,810$ & $49\,128\,815$ & $26\,587$ & $3\,406$ & $\mathbf{1\,580}$ \\
    \end{tabular}
    \end{minipage}
